---
title: "Homework 10"
output:
  pdf_document: default
  html_notebook: default
---

Homework	10	by Sana Khan:	I	consulted	with the textbook to complete the homework 


2. Download and library the nlme package and use data (“Blackmore”) to activate the Blackmore data set. Inspect the data and create a box plot showing the exercise level at different ages. Run a repeated measures ANOVA to compare exercise levels at ages 8, 10, and 12 using aov(). 


```{r}
 
#install.packages("nlme")
library(nlme)
library(car)
data("Blackmore")
summary(Blackmore)

Blackmore$age = round(Blackmore$age)

#boxplot

boxplot(exercise~age, data = Blackmore)

black_filtered <- Blackmore[Blackmore$age <= 12,]
black_filtered$age <- factor(black_filtered$age)

table(black_filtered$subject, black_filtered$age)


list <- rowSums(table(black_filtered$subject, black_filtered$age))==3
list <- list[list==TRUE]
list <- factor(names(list))
black_filtered <- black_filtered[black_filtered$subject %in% list,]
table(black_filtered$age)


summary(aov(exercise ~ age + Error(subject), data=black_filtered))


```

Based on the boxplot that was created, we can see that the median for the amount of exercise per age group is close to one, with variance increasing as age increases. The p-value we have is also statistically significant, which means we can reject the null value and say that as one ages the amount of time spent exercsing increases. 



5. Given that the AirPassengers data set has a substantial growth trend, use diff() to cre- ate a differenced data set. Use plot() to examine and interpret the results of differenc- ing. Use cpt.var() to find the change point in the variability of the differenced time series. Plot the result and describe in your own words what the change point signifies.

```{r}
library(changepoint)

data("AirPassengers")
diff_air <- diff(AirPassengers)

plot(diff_air)

cpt.var(diff_air)

plot(cpt.var(diff_air))

```
The command cpt.var calculates us when there is a changepoint in our dataset. The plot above shows the red line as the changepoint, which is sometime in 1955. 


Use cpt.mean() on the AirPassengers time series. Plot and interpret the results. Com- pare the change point of the mean that you uncovered in this case to the change point in the variance that you uncovered in Exercise 5. What do these change points suggest about the history of air travel?

```{r}
airmean=cpt.mean(AirPassengers)
airmean
plot(airmean)

airmean2 <- cpt.mean(AirPassengers, class=FALSE)

airmean2["conf.value"]
```

The plot shows us that there has been a shift in the mean over time, which was sometime in 1955. 

7.Find historical information about air travel on the Internet and/or in reference materials that sheds light on the results from Exercises 5 and 6. Write a mini-article (less than 250 words) that interprets your statistical findings from Exercises 5 and 6 in the context of the historical information you found.

According to https://theconversation.com/longing-for-the-golden-age-of-air-travel-be-careful-what-you-wish-for-34177, jets were intorduced in 1950, which meant it was faster to travel then before. In the 50's and 60's it become a bit more affordable to travel, with the introduction of new fares. The demographics of who was flying also changed, with more women and younger people flying. 


9.Use bcp() on the AirPassengers time series. Plot and interpret the results. Make sure to contrast these results with those from Exercise 6.

```{r}
#install.packages("bcp")
library(bcp)


bcpAir <- bcp(as.vector(AirPassengers), mcmc=1000, p0=1)
summary(bcpAir)


plot(bcpAir) 
plot(bcpAir$posterior.prob >.95,
     main="Plot of Air Passenger Posterior Probabilities > 95%",
     ylab="Posterior Probability", col="darkblue") 
```

The bayesian results are simliar to the frequentist method, where we can see a change point around 78. 