---
title: "Homework 5"
author: "Sana Khan"
date: "2022-11-06"
output: html_document
---

Homework 1 by Sana Khan: I consulted with our textbook to complete these problems.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

6**.The PlantGrowth data set contains three different groups, with each representing various plant food diets (you may need to type data(PlantGrowth) to activate it). The group labeled "ctrl" is the control group, while "trt1" and "trt2" are different types of experimental treatment. As a reminder, this subsetting statement accesses the weight data for the control group:**

**PlantGrowth**$weight[PlantGrowth$**group=="ctrl"] and this subsetting statement accesses the weight data for treatment group 1: PlantGrowth**$weight[PlantGrowth$**group=="trt1"]**

**Run a t‐test to compare the means of the control group ("ctrl") and treatment group 1 ("trt1") in the PlantGrowth data. Report the observed value of t, the degrees of freedom, and the p‐value associated with the observed value. Assuming an alpha threshold of .05, decide whether you should reject the null hypothesis or fail to reject the null hypothesis. In addition, report the upper and lower bound of the confidence interval.**

```{r pressure, echo=FALSE}
data(PlantGrowth)
ctrl <- PlantGrowth$weight[PlantGrowth$group=='ctrl']
trt1 <- PlantGrowth$weight[PlantGrowth$group=='trt1']

t.test(ctrl,trt1)

```

The value for T = 1.19, the degrees of freedom = 16.5 and the p-value =.25. The p-value is greater than our alpha, so we fail to reject the null hypothesis. The lower bound of our confidence interval is -.29 and the upper bound is 1.03

7**.Install and library() the BEST package. Note that you may need to install a program called JAGS onto your computer before you try to install the BEST package inside of R. Use BESTmcmc() to compare the PlantGrowth control group ("ctrl") to treatment group 1 ("trt1"). Plot the result and document the boundary values that BESTmcmc() calculated for the HDI. Write a brief definition of the meaning of the HDI and interpret the results from this comparison.**

```{r pressure, echo=FALSE}

plantBEST <- BESTmcmc(ctrl,trt1)

plantBEST

plot(plantBEST)


```

The boundary values for BESTmcmc are -.358 and 1.14. The HDI means that there is a 95% probability that the population means difference lies within the range of .358 and 1.14. There is an 86% chance that the population mean difference is greater than 0, which also supports failing to reject the null hypothesis.

8.**Compare and contrast the results of Exercise 6 and Exercise 7. You have three types of evidence: the results of the null hypothesis test, the confidence interval, and the HDI from the BESTmcmc() procedure. Each one adds something, in turn, to the understanding of the difference between groups. Explain what information each test provides about the comparison of the control group ("ctrl") and the treatment group 1 ("trt1").**

The three types of evidence we have show that the treatment group is not as effective as the control group. There isn't much of a difference in the results. The t-test shows us that the means are not very different and the p-value is greater than our alpha. The BESTmcmc also shows us that our range crosses over 0 with the majority falling below 0 which corresponds to failing to reject the null hypothesis.

9**.Using the same PlantGrowth data set, compare the "ctrl" group to the "trt2" group. Use all of the methods described earlier (t‐test, confidence interval, and Bayesian method) and explain all of the results**.

```{r}
trt2 <- PlantGrowth$weight[PlantGrowth$group=='trt2']

t.test(ctrl,trt2)

plantBEST2 <- BESTmcmc(ctrl,trt2)

plantBEST2

plot(plantBEST2)


```

The value for T = 2.14, the degrees of freedom = 16.78 and the p-value =.045. The p-value is less than our alpha, so we reject the null hypothesis. The lower bound of our confidence interval is -.0.983 and the upper bound is -.005. The boundary values for BESTmcmc are -1.04 and .08. The HDI means that there is a 95% probability that the population means difference lies within the range of -1.04 and .08. There is an 96% chance that the population mean difference is less than 0, which also supports rejecting the null hypothesis.

10. **Consider this t‐test, which compares two groups of n = 100,000 observations each:**

**t.test(rnorm(100000,mean=17.1,sd=3.8),rnorm(100000,mean=17.2,sd=3.8))**

**For each of the groups, the rnorm() command was used to generate a random normal distribution of observations similar to those for the automatic transmission group in the mtcars database (compare the programmed standard deviation for the random normal data to the actual mtcars data). The only difference between the two groups is that in the first rnorm() call, the mean is set to 17.1 mpg and in the second it is set to 17.2 mpg. I think you would agree that this is a negligible difference, if we are discuss‐ ing fuel economy. Run this line of code and comment on the results of the t‐test. What are the implications in terms of using the NHST on very large data sets?**

**t = -6.2855, df = 2e+05, p-value = 3.274e-10 alternative hypothesis: true difference in means is not equal to 0 95 percent confidence interval: -0.14002939 -0.07345877 sample estimates: mean of x mean of y 17.09753 17.20428**

The NHST can be unreliable on large data sets because small changes in the data can impact the results of rejecting or failing to reject the null hypothesis.
